---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# convoy-donations

<!-- badges: start -->
<!-- badges: end -->

This repository includes data and code to analyze donations made in support to the convoy that occupied Ottawa during the winter of 2022 using the Christian crowdsourcing platform GiveSendGo.

Load packages used for the analysis:
```{r}
library(cartogram)
library(tidyverse)
library(sf)
```

Read data:
```{r}
load(paste0(here::here(), "/data/da_2016_p1.rda"))
load(paste0(here::here(), "/data/da_2016_p2.rda"))
da_2016 <- rbind(da_2016_p1 %>% st_sf(),
                 da_2016_p2 %>% st_sf())
```

Filter data to investigate only the Hamilton CMA (CMA_UID == "35537"):
```{r}
hamilton_da_2016 <- da_2016 %>%
  filter(CMA_UID == "35537")
```

Project the geometry:
```{r}
hamilton_da_2016 <- hamilton_da_2016 %>%
  st_set_crs(4326) %>%
  st_transform(crs = 26917)
```

Plot geometry:
```{r}
ggplot() + 
  geom_sf(data = hamilton_da_2016)
```

Create population-based cartogram:
```{r}
DA_pop_cartogram <- cartogram_cont(hamilton_da_2016, 
                                   weight = "Population")
```

Summary:
```{r}
summary(hamilton_da_2016)
```

Replace NAs (donation_n and donation_amount):
```{r}
hamilton_da_2016 <- hamilton_da_2016 %>%
  mutate(donation_n = replace_na(donation_n, 0),
         donation_amount = replace_na(donation_amount, 0))
```

Total amount donated by residents of Hamilton CMA:
```{r}
sum(hamilton_da_2016$donation_amount)
```

Total number of donations by residents of Hamilton CMA:
```{r}
sum(hamilton_da_2016$donation_n)
```

Number of individual donations divided by population:
```{r}
sum(hamilton_da_2016$donation_n)/sum(hamilton_da_2016$Population) * 100
```

Number of DAs with at least one donation:
```{r}
hamilton_da_2016 %>%
  filter(donation_n != 0) %>%
  nrow()
```



Donations:
```{r}
ggplot(hamilton_da_2016,
       aes(x = donation_amount)) +
  geom_histogram(binwidth = 50) + 
  xlab("Donation amount by DA ($)") + 
  ylab("Count of Dissemination Areas") +
  theme_minimal()
```


Plot number of donations:
```{r}
ggplot() + 
  geom_sf(data = hamilton_da_2016 %>%
            mutate(donation_n = case_when(donation_n == 0 ~ "0",
                                          donation_n > 0 & donation_n <= 5 ~ "(0, 5]",
                                          donation_n > 5 & donation_n <= 10 ~ "(5, 10]",
                                          donation_n > 10 ~ "(10, 15]"),
                   donation_n = factor(donation_n, 
                                       levels = c("0", "(0, 5]", "(5, 10]", "(10, 15]"),
                                       ordered = TRUE)),
          aes(fill = donation_n),
          color = "black",
          size = 0.01) +
  scale_fill_manual(name = "Individual donations",
                    values = c("0" = "white", 
                               "(0, 5]" = "darkgoldenrod1", 
                               "(5, 10]" = "darkorange", 
                               "(10, 15]" = "brown1")) +
  theme_minimal()
```

Plot number of donations in the cartogram:
```{r}
ggplot() + 
  geom_sf(data = DA_pop_cartogram %>%
            mutate(donation_n = case_when(donation_n == 0 ~ "0",
                                          donation_n > 0 & donation_n <= 5 ~ "(0, 5]",
                                          donation_n > 5 & donation_n <= 10 ~ "(5, 10]",
                                          donation_n > 10 ~ "(10, 15]"),
                   donation_n = factor(donation_n, 
                                       levels = c("0", "(0, 5]", "(5, 10]", "(10, 15]"),
                                       ordered = TRUE)),
          aes(fill = donation_n),
          color = "black",
          size = 0.01) +
  scale_fill_manual(name = "Individual donations",
                    values = c("0" = "white", 
                               "(0, 5]" = "darkgoldenrod1", 
                               "(5, 10]" = "darkorange", 
                               "(10, 15]" = "brown1")) +
  theme_minimal()
```

Plot amount of donations:
```{r}
ggplot() + 
  geom_sf(data = hamilton_da_2016 %>%
            mutate(donation_amount = case_when(donation_amount == 0 ~ "0",
                                          donation_amount > 0 & donation_amount <= 50 ~ "(0, 50]",
                                          donation_amount > 50 & donation_amount <= 100 ~ "(50, 100]",
                                          donation_amount > 100 ~ ">100"),
                   donation_amount = factor(donation_amount, 
                                       levels = c("0", "(0, 50]", "(50, 100]", ">100"),
                                       ordered = TRUE)),
          aes(fill = donation_amount),
          color = "black",
          size = 0.01) +
  scale_fill_manual(name = "Total amounts by DA",
                    values = c("0" = "white", 
                               "(0, 50]" = "darkgoldenrod1", 
                               "(50, 100]" = "darkorange", 
                               ">100" = "brown1")) +
  theme_minimal()
```

Plot number of donations in the cartogram:
```{r}
ggplot() + 
  geom_sf(data = DA_pop_cartogram %>%
            mutate(donation_amount = case_when(donation_amount == 0 ~ "0",
                                          donation_amount > 0 & donation_amount <= 50 ~ "(0, 50]",
                                          donation_amount > 50 & donation_amount <= 100 ~ "(50, 100]",
                                          donation_amount > 100 ~ ">100"),
                   donation_amount = factor(donation_amount, 
                                       levels = c("0", "(0, 50]", "(50, 100]", ">100"),
                                       ordered = TRUE)),
          aes(fill = donation_amount),
          color = "black",
          size = 0.01) +
  scale_fill_manual(name = "Total amounts by DA",
                    values = c("0" = "white", 
                               "(0, 50]" = "darkgoldenrod1", 
                               "(50, 100]" = "darkorange", 
                               ">100" = "brown1")) +
  theme_minimal()
```
